---
title: Lab 5
output: pdf_document
date: "2023-03-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
install.packages("rdrobust")
library(rdrobust)
library(lmtest)
df = read_dta("probation.dta")
view(df)
summary(df)
```

***Question 1***
We don't want to  all students who are on probation with all students who are not on probation to evaluate this University’s program because we are trying evaluate the potential causal impact of the probation policy on educational outcomes. To do so, we are employing a regression discontuinity design that examines educational outcomes of students who did and did not receive probation but are otherwise roughly the same in other characteristics. This requires us to only look at observations right around (ie, right above and below) the cutoff of the 1.60 GPA threshold that triggers a student receiving academic probation.

***Question 2***
The running variable is the student's GPA. The cutoff is whether that GPA is below or above 1.6

```{r q3}


#Question 3



```





```{r}

#q1 first create the distnace from the GPA cutoff
df$gpa_cutoff = df$GPA - 1.6
df_narrow = df |> filter(gpa_cutoff > 1.2 & gpa_cutoff <1.2)

rdplot(x = df_narrow$hsgrade_pct,
       y = df_narrow$gradin4,
       c = 0,
       p = 1 
       x.label = "GPA Cutoff",
       y = "Age at Freshman Year"
       
       
       #plot a histogram for thsi cutoff
       #here we would like to udnerstand wherehter there is bunching at the cutoff or if there is a jump in the # of students or just about the same
       
       
       ggplot(data = df_narrow, aes(x = GPA, y= ..density..))+
         geom_histogram(bins = 400)+
         geom_vline(xintercept = 1.6, color = 'red')
       
      #now do for the centered cutoff
       ggplot(df_narrow, aes(x = gpa_cutoff, y = ..density..))+
         geom_histogram(bins = 400) + geom_vline(xintercept = 0)
       
       
       #question 4
       prediction2 = reg2$coefficients[1] + reg2$coeffienceits[2]*1.6
       
       #question 7
       df_narrow$cutoff = ifelse(dfnarrow$gpa >= 1.6, 1, 0 )
       ilai_reg = lm(grain4 ~ cutoff +gpa_cutoff + cutoff*gpa_cutoff, data = df_narrow
                     coeftest(ilai_reg, vcov = vcovHC(ilai_reg, type = "HC1"))
                     
                     df_narrow$cutoff = ifelse(df_narrow$gpa > 1.6, 1, 0)
                     reg3 = lm(gradin4 ~ cutoff + gpa_cutoff + cutoff*gpa_cutoff, data = df_narrow)
                     coeftest(reg3, vcov = vcovHC(Reg3, type = "HC1")) 
       
       



```

